```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "Interaction Report: NT FvF Dataset"
author: "Michael Cowled"
output:
  html_document:
    df_print: paged
---
```{r, echo = FALSE}
paste0("Date is: ", format(Sys.time(), "%a %d %b %Y"))
```

## R Packages required to be loaded in:

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(data.table)
library(ggplot2)
library(ggdendro)
library(reshape2)
library(readxl)
```

## Heatmap showing the average effect on the metabolome of each interaction

Note: Ignores both inductions and complete suppressions since the corresponding metabolites are not present in both monoculture and coculture.

```{r, echo = FALSE}
filenames <- list.files(path = "Testing Broad-Scale Interactions/OutputFiles/NT_FvF", pattern = "F", full.names = TRUE)
my.data <- lapply(filenames, read.csv)
full.list <- rbindlist(my.data, use.names=TRUE, fill=FALSE)

#For changing NT to F in NT_FvA
full.list$Sample_Ref <- gsub("NT", "F", full.list$Sample_Ref, perl = TRUE)
full.list$Sample_Ref <- gsub("F17", "B1", full.list$Sample_Ref, perl = TRUE)
full.list$Sample_Ref <- gsub("F18", "B2", full.list$Sample_Ref, perl = TRUE)
full.list$Sample_Ref <- gsub("F19", "B3", full.list$Sample_Ref, perl = TRUE)

refined.list <- filter(full.list, Metabolite_Effect < 5 & Metabolite_Effect > 1)
cond.5.list <- filter(full.list, Metabolite_Effect == 5)
cond.5.list$PeakRatio <- 100
refined.list <- rbind(refined.list, cond.5.list)
Cocultures <- c(unique(refined.list$Sample_Ref))

ratio.df <- data.frame(matrix(ncol = 2, nrow = 0))
for (i in Cocultures[1:length(Cocultures)]) {
    temp <- filter(refined.list, Sample_Ref == i)
    ratio.df <- rbind(ratio.df, 
                      c(Sample_Ref = i, Avg_Ratio = mean(temp$PeakRatio, na.rm = TRUE)))
}

names(ratio.df) <- c("Sample_Ref", "Avg_Ratio")
ratio.df <- separate(ratio.df, Sample_Ref, into = c("Ref_Culture", "Int_Culture"), sep = "v")
ratio.df$Avg_Ratio <- as.numeric(ratio.df$Avg_Ratio)

#For NT fungi
ratio.df$Ref_Culture <- gsub("(?<![0-9])([0-9])(?![0-9])", "0\\1", ratio.df$Ref_Culture, perl = TRUE)
ratio.df$Int_Culture <- gsub("(?<![0-9])([0-9])(?![0-9])", "0\\1", ratio.df$Int_Culture, perl = TRUE)
Fungi_Assignments <- read_excel("Testing Broad-Scale Interactions/Fungi_Assignments_nt.xlsx")
ratio.df2 <- merge(ratio.df, Fungi_Assignments, by="Ref_Culture", all.x=TRUE)
Fungi_Assignments <- read_excel("Testing Broad-Scale Interactions/Fungi_Assignments_nt2.xlsx")
ratio.df3 <- merge(ratio.df2, Fungi_Assignments, by="Int_Culture", all.x=TRUE)

heatmap.fvf.avg <- ggplot(data = ratio.df3, aes(x=Int_Species, y=Ref_Species)) + 
    geom_tile(aes(fill=Avg_Ratio)) +
    scale_fill_gradient2(low = "dark blue", high = "dark red", mid = "white", 
                         midpoint = 0, space = "Lab", name="Ratio (%)") +
    labs(x="Interacting Culture", y="Reference Culture") +
    theme_grey(base_size=8) +
    theme(axis.text.x=element_text(angle=45, hjust = 1))
heatmap.fvf.avg
ggsave(heatmap.fvf.avg,filename="heatmap.fvf.avg.png",height=2.24,width=3,units="in",dpi=200)
```

## Heatmaps showing the effect of each interaction on a particular fungus

```{r, echo = FALSE, warning = FALSE}
coculture.list <- filter(full.list, !is.na(PeakNo_CC)) %>%
    separate(Sample_Ref, into = c("Ref_Culture", "Int_Culture"), sep = "v")
coculture.list$Ref_Culture <- gsub("(?<![0-9])([0-9])(?![0-9])", "0\\1", coculture.list$Ref_Culture, perl = TRUE)
coculture.list$Int_Culture <- gsub("(?<![0-9])([0-9])(?![0-9])", "0\\1", coculture.list$Int_Culture, perl = TRUE)
coculture.list$Matched_con <- gsub("(?<![0-9])([0-9])(?![0-9])", "0\\1", coculture.list$Matched_con, perl = TRUE)
coculture.list <- arrange(coculture.list, Ref_Culture)
Fungi_Assignments <- read_excel("Testing Broad-Scale Interactions/Fungi_Assignments_nt.xlsx")
coculture.list <- merge(coculture.list, Fungi_Assignments, by="Ref_Culture", all.x=TRUE)
Fungi_Assignments <- read_excel("Testing Broad-Scale Interactions/Fungi_Assignments_nt2.xlsx")
coculture.list <- merge(coculture.list, Fungi_Assignments, by="Int_Culture", all.x=TRUE)
Fungi_Assignments <- read_excel("Testing Broad-Scale Interactions/Fungi_Assignments_nt3.xlsx")
coculture.list <- merge(coculture.list, Fungi_Assignments, by="Matched_con", all.x=TRUE)
fungus <- unique(coculture.list$Ref_Species)

for (i in fungus[1:length(fungus)]) {
subset.list <- filter(coculture.list, Ref_Species == i) %>%
  filter(Matched_Species == paste0(i, "CON") | is.na(Matched_Species)) %>%
    select(Int_Species, PeakNo_con, PeakRatio) %>%
    mutate(logPeakRatio = log((PeakRatio/100) + 1))
           
hm.title <- i
heatmap_fvf <- ggplot(data = subset.list, aes(x=PeakNo_con, y=Int_Species)) + 
    geom_tile(aes(fill=logPeakRatio)) +
    scale_fill_gradient2(low = "dark blue", high = "dark red", mid = "white", 
                         midpoint = 0, limit = c(min(subset.list$logPeakRatio), 
                                                 max(subset.list$logPeakRatio)), 
                         space = "Lab", name="log(%PeakArea)") +
    labs(x="Peak Number", y="Coculture") +
    theme_grey(base_size=8) +
    ggtitle(label = paste0("Interactions involving ", i))
print(heatmap_fvf)
ggsave(heatmap_fvf,filename=paste0("heatmap_fvf", i, ".png"),height=1.75,width=10.20,units="in",dpi=200)
}
```

## Barchart showing the proportion of effects based on retention time for each fungus

```{r, echo = FALSE, warning = FALSE}
mean.list <- setnames(data.frame(matrix(ncol = 8, nrow = 15)), c("mE5", "mE6", "mE5&6", "mAll",
                                                                 "eE5", "eE6", "eE5&6", "eAll"))

for (i in 1:15) {
x <- paste0("NT", i, "v")
filenames <- list.files(path = "Testing Broad-Scale Interactions/OutputFiles/NT_FvA", pattern = x, full.names = TRUE)
my.data <- lapply(filenames, read.csv)
full.list <- rbindlist(my.data, use.names=TRUE, fill=FALSE)
full.list <- transform(full.list, Metabolite_Effect = factor(Metabolite_Effect))
full.list <- filter(full.list, Metabolite_Effect != 1)

Effect_5 <- filter(full.list, Metabolite_Effect == 5)
Effect_6 <- filter(full.list, Metabolite_Effect == 6)
Effect_5_6 <- filter(full.list, as.numeric(Metabolite_Effect) > 4)

mean.list[i, 1] <- mean(Effect_5$RetTime_CC)
mean.list[i, 2] <- mean(Effect_6$RetTime_CC)
mean.list[i, 3] <- mean(Effect_5_6$RetTime_CC)
mean.list[i, 4] <- mean(full.list$RetTime_CC)
mean.list[i, 5] <- error <- qnorm(.95)*(sd(Effect_5$RetTime_CC)/sqrt(length(Effect_5$RetTime_CC)))
mean.list[i, 6] <- error <- qnorm(.95)*(sd(Effect_6$RetTime_CC)/sqrt(length(Effect_6$RetTime_CC)))
mean.list[i, 7] <- error <- qnorm(.95)*(sd(Effect_5_6$RetTime_CC)/sqrt(length(Effect_5_6$RetTime_CC)))
mean.list[i, 8] <- error <- qnorm(.95)*(sd(full.list$RetTime_CC)/sqrt(length(full.list$RetTime_CC)))

}

for (i in 1:15) {
    col1 <- c(paste0("F", i), paste0("F", i), paste0("F", i), paste0("F", i))
    if (i > 1) {
        extraction2 <- cbind(col1, c("E5", "E6", "E5 & E6", "All"), 
                                              transpose(mean.list[i,1:4]), transpose(mean.list[i,5:8]))
        names(extraction2) <- names(extraction)
        extraction <- rbind(extraction, extraction2)
    } else {
        extraction <- cbind(col1, c("E5", "E6", "E5 & E6", "All"), 
                            transpose(mean.list[i,1:4]), transpose(mean.list[i,5:8]))
        names(extraction) <- c("fungus", "Effect", "mean", "se")
        }
}

extraction$fungus <- gsub("(?<![0-9])([0-9])(?![0-9])", "0\\1", extraction$fungus, perl = TRUE)
Fungi_Assignments <- read_excel("Testing Broad-Scale Interactions/Fungi_Assignments_nt4.xlsx")
extraction <- merge(extraction, Fungi_Assignments, by="fungus", all.x=TRUE)

p<- ggplot(extraction, aes(x=Ref_Species, y=mean, fill=Effect)) + 
    geom_bar(stat="identity", color="black", 
             position=position_dodge()) +
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.4,
                  position=position_dodge(.9)) 
barchart <- p+labs(x="Reference Fungus", y = "Retention Time")+
    theme_classic() +
    theme(axis.text.x=element_text(angle=45, hjust = 1))
barchart
ggsave(barchart,filename="barchart_fvf_tal.png",height=3.5,width=10.20,units="in",dpi=200)
```

## Boxplots showing the overall effect on retention time for each metabolite effect category (and associated histograms)

```{r, echo = FALSE, warning = FALSE}
filenames <- list.files(path = "Testing Broad-Scale Interactions/OutputFiles/NT_FvA", pattern = "NT", full.names = TRUE)
my.data <- lapply(filenames, read.csv)
full.list <- rbindlist(my.data, use.names=TRUE, fill=FALSE)
full.list$Metabolite_Effect <- as.numeric(full.list$Metabolite_Effect)

#Filters into different lists based on metabolite_effect
Effect_1 <- filter(full.list, as.numeric(Metabolite_Effect) == 1)
Effect_2 <- filter(full.list, as.numeric(Metabolite_Effect) == 2)
Effect_3 <- filter(full.list, as.numeric(Metabolite_Effect) == 3)
Effect_4 <- filter(full.list, as.numeric(Metabolite_Effect) == 4)
Effect_5 <- filter(full.list, as.numeric(Metabolite_Effect) == 5)
Effect_6 <- filter(full.list, as.numeric(Metabolite_Effect) == 6)
All_Effects_Less_Induction <- filter(full.list, as.numeric(Metabolite_Effect) < 5 & as.numeric(Metabolite_Effect) > 1)
Effect_5_6 <- filter(full.list, as.numeric(Metabolite_Effect) > 4)

full.list <- transform(full.list, Metabolite_Effect = factor(Metabolite_Effect))
full.list.minus.effect.1 <- filter(full.list, as.numeric(Metabolite_Effect) <7 & as.numeric(Metabolite_Effect) > 1)
effect.names <- c("Suppression", "Little to no change", "Enhancement", "Major enhancement", "Induction")
ggplot(data=full.list.minus.effect.1, aes(x=Metabolite_Effect, y=RetTime_CC)) + geom_boxplot(aes(fill=Metabolite_Effect)) + 
    ylab("Retention Time (min)") + xlab("Metabolite Effect") +
    stat_summary(fun.y=mean, geom="point", shape=1, size=4) +
    theme(axis.text.x = element_text(angle = 90, face = "italic")) + scale_x_discrete(labels = effect.names)

par(mfrow=c(2,3))
hist(Effect_1$RetTime_con, main = "Complete Suppression", xlab = "Retention Time", ylab = "Number of Secondary Metabolites")
hist(Effect_2$RetTime_CC, main = "Suppression", xlab = "Retention Time", ylab = "Number of Secondary Metabolites")
hist(Effect_3$RetTime_CC, main = "Little to No Change", xlab = "Retention Time", ylab = "Number of Secondary Metabolites")
hist(Effect_4$RetTime_CC, main = "Enhancement", xlab = "Retention Time", ylab = "Number of Secondary Metabolites")
hist(Effect_5$RetTime_CC, main = "Major Enhancement", xlab = "Retention Time", ylab = "Number of Secondary Metabolites")
hist(Effect_6$RetTime_CC, main = "Induction or Unmatched", xlab = "Retention Time", ylab = "Number of Secondary Metabolites")
hist(All_Effects_Less_Induction$RetTime_CC, main = "All other effects", xlab = "Retention Time", ylab = "Number of Secondary Metabolites")
hist(Effect_5_6$RetTime_CC, main = "Induction or Major Enhancement", xlab = "Retention Time", ylab = "Number of Secondary Metabolites")
```