```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "MIMI: Microbial Interaction Metabolite Integrator"
author: "Michael Cowled"
output:
  html_document:
    df_print: paged
---
```{r, echo = FALSE}
paste0("Date is: ", format(Sys.time(), "%a %d %b %Y"))
```

## R Packages required to be loaded in:

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(readxl)
```

## NovaC files

NovaC is an unpublished data extraction and formatting tool for Agilent Chemstation LCMS/HPLC files owned by Macquarie University and created by A. Prof. Andrew Piggott (Molecular Sciences/Faculty of Science & Engineering).

The raw data was extracted from Agilent Chemstation LCMS/HPLC using the macro "NovaC.mac" (Macquarie Unviersity) which generates an Excel file with the following name: "002-D1F-D7-_F7vF2.D_NovaC", as an example.

See example NovaC files in folder titled: "Example_NovaC_Data".

File names were manually tidied up to remove unnecessary prefixes or suffixes: "F7vF2".

This name tidying and the creation of the Interaction_Matrix below, are the only manual steps performed.

## The Interaction Matrix

A table in excel format with the heading names as follows is required to be loaded in by the user.
The matrix should contain all interactions to be investigated.

```{r}
Interaction_Matrix <- read_excel("Interaction_Matrix.xlsx")
head(Interaction_Matrix)
```

## MIMI - Script #1/2

The following functions are to be pre-loaded prior to use of the main function, MIMI()

1. **ReadExcel:** Reads in the NovaC files corresponding to a particular row number in the Interaction_Matrix.
2. **CheckUVCount:** Compares the top 5 UV maxima of a matched peak in the control of interest and the coculture.
3. **ReadUV:** Reads in the UV spectral data (abs vs. wavelength) for the corresponding 3 samples being compared.
4. **SubtractUV:** Subtracts the UV spectrum of the control of interest from the coculture, and takes the mean(Abs).
5. **RowBinder:** Binds the matched peak found in PeakMatcher to a df named cc.df.
6. **CalcRatio:** Calculates the %Enhancement/Suppresion compared to control levels.
7. **PeakMatcher:** Matches and verifies peaks from the control of interest to peaks in the coculture, utilising a combination of retention time, number of matching UV maxima (UVcheck) and the means of the subtracted UV spectra (UVsubtract).
8. **ConConsolidator:** Consolidates the outcome table to match peaks from coculture to a single peak in a control. Compares subtracted UV spectra to make decisions based on double matching.
9. **EffectCategoriser:** Characterises the Peak Area ratio as an effect to the metabolite in the coculture (induction, suppression, etc.).

## MIMI - Script #2/2

The following functions are to be pre-loaded prior to use of the main function, MIMI2()

1. **SimpleEffectCharacteriser:** Uses the principles of the EffectCategoriser function to categorise based on a single PeakRatio input.
2. **DoublePeakChecker:** Finds and reports instances of double peak matching.
3. **IdentifyBadPeak:** Decides on which of the doubley assigned peaks to remove.
4. **RemoveDoubleyAssignedPeaks:** Checks for a peak in the control being assigned to more than one peak in the coculture. Uses the number of matching UV maxima and/or the subtracted UV spectra to make decisions as to which peak is a better match.
5. **PeakAssigner:** Adds the assignment of the matched non UV peak from MatchNonUVs.
6. **MatchNonUVs:** Tentatively assigns matched peaks as non UVs (or as distorted UVs) if matching the conditions.
7. **RowBinder2:** A variant of RowBinder used for adding missing control peaks.
8. **FindMissingcontrolPeaks:** Adds in unassigned peaks from the controls to provide a single, unified table.

Please see the full script to load in the whole suite of functions and review annotations for extra details.
The final main functions, **MIMI()** and **MIMI2()**, combines the use of all 17 functions in providing the final tidied output files.

## The output file

An example output file is as follows:

```{r, echo = FALSE}
head(read.csv("Testing Broad-Scale Interactions/OutputFiles/NT_FvF/F1vF3.csv"))
```

The resulting output file contains the following columns containing:

1. **Sample_Ref:** The reference sample - the default is the coculture, unless the peak from a control is not present.
2. **PeakNo_CC:** The peak number as present in the coculture sample.
3. **RetTime_CC:** The retention time (in min) for the peak in the coculture sample.
4. **PeakArea_CC:** The peak area for the peak in the coculture sample.
5. **PercArea:** The percentage the peak area associated with this peak accounts for compared to the rest of the peaks in the sample.
6. **Matched_con:** The control sample with a matching peak to the sample reference.
7. **PeakNo_con:** The peak number as present in the matched control or reference control sample.
8. **RetTime_con:** The retention time (in min) for the peak in the control sample.
9. **PeakArea_con:** The peak area for the peak in the control sample.
10. **UV_Count:** The number of UV maxima that match between the control and coculture peaks.
11. **Subtracted_UV_Mean:** The mean of the subtracted UV spectra (in Absorbance units) of the control from the coculture.
12. **PeakRatio:** The relative ratio (in %) of the reference sample compared to the control.
13. **Metabolite_Effect:** A categorised effect on the metabolite in the reference sample as determined by the PeakRatio.
These effects are:
*Complete Suppression:* Peak Area = -100%
| *Suppression:* -100% < Peak Area <= -20%
| *Little to No Change:* -20% < Peak Area < 20%
*Enhancement:* 20 <= Peak Area < 100%
| *Major enhancement:* Peak area >=100%
| *Induction:* NA (No peak area to compare to in CON)