```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "Interaction Report: NT FvF Dataset"
author: "Michael Cowled"
output:
  html_document:
    df_print: paged
---
```{r, echo = FALSE}
paste0("Date is: ", format(Sys.time(), "%a %d %b %Y"))
```

## R Packages required to be loaded in:

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(data.table)
library(ggplot2)
library(ggdendro)
library(reshape2)
library(readxl)
library(ggtext)
library(glue)
library(gplots)
```

## Heatmap showing the average effect on the metabolome of each interaction

Note: Ignores both inductions and complete suppressions since the corresponding metabolites are not present in both monoculture and coculture.

```{r, echo = FALSE}
filenames <- list.files(path = "Testing Broad-Scale Interactions/OutputFiles/AvNOVO", pattern = "", full.names = TRUE)
my.data <- lapply(filenames, read.csv)
full.list <- rbindlist(my.data, use.names=TRUE, fill=FALSE)

#For changing NT to F in NT_FvA
full.list$Sample_Ref <- gsub("NT", "F", full.list$Sample_Ref, perl = TRUE)
full.list$Sample_Ref <- gsub("F17", "B1", full.list$Sample_Ref, perl = TRUE)
full.list$Sample_Ref <- gsub("F18", "B2", full.list$Sample_Ref, perl = TRUE)
full.list$Sample_Ref <- gsub("F19", "B3", full.list$Sample_Ref, perl = TRUE)

refined.list <- filter(full.list, Metabolite_Effect < 5 & Metabolite_Effect > 1)
cond.5.list <- filter(full.list, Metabolite_Effect == 5)
cond.5.list$PeakRatio <- 100
refined.list <- rbind(refined.list, cond.5.list)
Cocultures <- c(unique(refined.list$Sample_Ref))

ratio.df <- data.frame(matrix(ncol = 2, nrow = 0))
for (i in Cocultures[1:length(Cocultures)]) {
    temp <- filter(refined.list, Sample_Ref == i)
    ratio.df <- rbind(ratio.df, 
                      c(Sample_Ref = i, Avg_Ratio = mean(temp$PeakRatio, na.rm = TRUE)))
}

names(ratio.df) <- c("Sample_Ref", "Avg_Ratio")
ratio.df <- separate(ratio.df, Sample_Ref, into = c("Ref_Culture", "Int_Culture"), sep = "_")
ratio.df$Avg_Ratio <- as.numeric(ratio.df$Avg_Ratio)

#For NT fungi
ratio.df$Ref_Culture <- gsub("(?<![0-9])([0-9])(?![0-9])", "0\\1", ratio.df$Ref_Culture, perl = TRUE)
ratio.df$Int_Culture <- gsub("(?<![0-9])([0-9])(?![0-9])", "0\\1", ratio.df$Int_Culture, perl = TRUE)
ratio.df <- mutate(ratio.df, Int_Culture_2 = "Novobiocin")

color <- "grey30"
Actino_Assignments <- read_excel("Testing Broad-Scale Interactions/Actino_Assignments.xlsx")
ratio.list <- merge(ratio.df, Actino_Assignments, by="Ref_Culture", all.x=TRUE)
ratio.list <- mutate(ratio.list, Species = glue("<i style='color:{color}'>{Sp1}</i> {Sp2}"))
condensed.list <- select(ratio.list, c("Species", "T2PKS", "PPTS Gene", "Avg_Ratio"))
long <- melt(condensed.list, id.vars = c("Species"))

heatmap.fvf.avg <- ggplot(data = long, aes(x=variable, y=Species)) + 
    geom_tile(aes(fill=value)) +
    scale_fill_gradient2(low = "dark blue", high = "dark red", mid = "white", 
                         midpoint = 0, space = "Lab", name="Ratio (%)") +
    labs(x="Treatment", y="Reference Culture") +
    theme(axis.text.y = element_markdown())
heatmap.fvf.avg
ggsave(heatmap.fvf.avg,filename="heatmap.fvf.avg.png",height=2.24,width=3,units="in",dpi=200)
```

###testing (clustered heatmap)

```{r, echo = FALSE}
ratio.list <- merge(ratio.df, Actino_Assignments, by="Ref_Culture", all.x=TRUE)
ratio.list <- mutate(ratio.list, Species = glue("{Sp1} {Sp2}"))
df.as.matrix <- select(ratio.list, Species, T2PKS, "PPTS Gene", Avg_Ratio) 
names <- as.vector(df.as.matrix$Species)
df.as.matrix <- select(df.as.matrix, -Species)

df.as.matrix <- as.matrix(df.as.matrix)
row.names(df.as.matrix) <- names
colnames(df.as.matrix) <- c("T2PKS", "PPTS", "Ratio")

colfunc <- colorRampPalette(c("dark blue", "white", "dark red"))

heatmap.2(df.as.matrix, col=colfunc(25), scale = "none", trace = "none", cexCol=1, cexRow = 0.5, margin=c(5, 25))
```

## Heatmaps showing the effect of each interaction on a particular fungus

```{r, echo = FALSE, warning = FALSE}
coculture.list <- filter(full.list, !is.na(PeakNo_CC)) %>%
    separate(Sample_Ref, into = c("Ref_Culture", "Int_Culture"), sep = "_")
coculture.list$Ref_Culture <- gsub("(?<![0-9])([0-9])(?![0-9])", "0\\1", coculture.list$Ref_Culture, perl = TRUE)
coculture.list$Int_Culture <- gsub("(?<![0-9])([0-9])(?![0-9])", "0\\1", coculture.list$Int_Culture, perl = TRUE)
coculture.list$Matched_con <- gsub("(?<![0-9])([0-9])(?![0-9])", "0\\1", coculture.list$Matched_con, perl = TRUE)
coculture.list <- arrange(coculture.list, Ref_Culture)
fungus <- unique(coculture.list$Ref_Culture)

i <- 1

while (i <= length(fungus)) {
subset.list <- filter(coculture.list, Ref_Culture == fungus[i]) %>%
  filter(Matched_con == paste0(fungus[i], "_CON") | is.na(Matched_con)) %>%
    select(Int_Culture, Ref_Culture, PeakNo_con, PeakRatio, Metabolite_Effect) %>%
    mutate(logPeakRatio = log((PeakRatio/100) + 1)) %>%
    mutate(Int_Culture_2 = "")
           
hm.title <- fungus[i]
heatmap_fvf <- ggplot(data = subset.list, aes(x=PeakNo_con, y=Int_Culture_2)) + 
    geom_tile(aes(fill=PeakRatio)) +
    scale_fill_gradient2(low = "dark blue", high = "dark red", mid = "white", 
                         midpoint = 0, limit = c(min(subset.list$PeakRatio), 
                                                 max(subset.list$PeakRatio)), 
                         space = "Lab", name="Ratio (%)") +
    labs(x="Peak Number", y="Treatment") +
    theme_grey(base_size=8) +
    ggtitle(label = paste0("Effect on ", fungus[i]))
print(heatmap_fvf)
ggsave(heatmap_fvf,filename=paste0("heatmap_fvf", fungus[i], ".png"),height=1.25,width=10.20,units="in",dpi=200)

if (i > 1) {
    X.list <- rbind(X.list, subset.list)
}   else {
    X.list <- subset.list
}
i <- i + 1
}

X.list$PeakRatio[X.list$PeakRatio > 300] <- 300

write.csv(X.list, "Testing Broad-Scale Interactions/OutputFiles/vKit_list.CSV", row.names = FALSE)
color <- "grey30"
Actino_Assignments <- read_excel("Testing Broad-Scale Interactions/Actino_Assignments.xlsx")
Y.list <- merge(X.list, Actino_Assignments, by="Ref_Culture", all.x=TRUE)
Y.list <- mutate(Y.list, Species = glue("<i style='color:{color}'>{Sp1}</i> {Sp2}"))

heatmap_fvf <- ggplot(data = Y.list, aes(x=PeakNo_con, y=Species)) + 
    geom_tile(aes(fill=PeakRatio)) +
    scale_fill_gradient2(low = "dark blue", high = "dark red", mid = "white", 
                        midpoint = 0, limit = c(-100, max(Y.list$PeakRatio)), 
                       space = "Lab", name="Ratio (%)") +
    labs(x="Peak Number", y="Reference culture") +
    theme(axis.text.y = element_markdown())
print(heatmap_fvf)
ggsave(heatmap_fvf,filename=paste0("heatmap_fvf", i, ".png"),height=2.24,width=10.20,units="in",dpi=200)

```